Tomographic galaxy clustering with the Subaru Hyper Suprime-Cam first year public data release

### 2019-12-16 doc/Paper/main.tex
Added acknowledgments for Jo.
Started to incorporate comments from Eric.

  \begin{table}
  \centering
  \begin{tabular}{|l|l|l|l|}
    \hline
    $z_{\rm phot}^{\rm ini}$ & $z_{\rm phot}^{\rm end}$ & $\bar{z}$ & $N_{\rm gal}$ \\
    \hline
    0.15 & 0.5  & 0.57 & 1,750,274 \\
    0.5  & 0.75 & 0.68 & 1,766,939 \\
    0.75 & 1.0  & 0.91 & 1,702,685 \\
    1.0  & 1.5  & 1.26 & 1,752,359 \\
    \hline
  \end{tabular}
  \caption{Summary of the 4 redshift bins used in our analysis. The first two columns show the photo-$z$ bin edges. We used the {\tt photoz\_best} redshift estimator for the photo-$z$ code {\tt Ephor\_AB} to assign galaxies to different bins. The second column shows the mean redshift of each bin calculated from the fiducial redshift distributions described in Section \ref{ssec:methods.nz}. The last column shows the number of galaxies in each bin. The bin edges were chosen to roughly contain an equal number of galaxies in each of them.} \label{tab:bins_summary}
  \end{table}

@@

  \subsection{Redshift distributions}\label{ssec:methods.nz}
    \begin{figure}
      \centering
      \includegraphics[width=0.9\textwidth]{figures/nzs.pdf}
      \caption{Redshift distributions used in this analysis. The histograms in each panel are the same, and correspond to the redshift distributions estimated from the COSMOS 30-band catalog. These are the fiducial redshift distributions used in our analysis. The solid curves in each panel show the stacked redshift distributions inferred by stacking the per-object photo-$z$ probability distributions obtained with the four alternative photo-$z$ codes explored here ({\tt DEmP}, {\tt Ephor}, {\tt Ephor\_AB} and {\tt FRANKEN-Z} from left to right and top to bottom). The redshift distributions for the four different redshift bins listed in Table \ref{tab:bins_summary} are shown in different colors. Note that all four redshift estimation codes use the same COSMOS 30-band catalog so the level of visual agreement should not be taken as realistic measure of uncertainty on the redshift distributions.}
      \label{fig:nzs}
    \end{figure}
    The underlying redshift distributions of our tomographic samples ($p^i(z)$ in Eq. \ref{eq:cell_gg_limber}) are a central component of the theory model used to infer astrophysical and cosmological parameters. Estimating redshift distributions for photometric samples is a non-trivial problem that has been studied extensively in the literature \citep{2008MNRAS.390..118L,2008ApJ...684...88N,2018MNRAS.478..592H}. We use two different methods to estimate $p^i(z)$.

@@

      \paragraph{Number densities.} The number density of galaxies estimated in a given field is simply given by the ratio of the number of galaxies and the area of the field:
      \begin{equation}
        \hat{\bar{n}}=\frac{\sum_p N_p}{\sum_p \Omega_p}= \bar{n} \frac{\int \mathrm{d}\nv^2\,W(\nv)\left[1+\delta_g(\nv)\right]}{\int \mathrm{d}\nv^2\,W(\nv)},
      \end{equation}
      where $p$ runs over all pixels in the map, and $N_p$ and $\Omega_p$ are the number of objects and area of pixel $p$. In the second equality we have taken the continuum limit, $\bar{n}$ is the true number density, $\delta_g$ is the galaxy overdensity and $W$ is the field's mask. Using the statistics of $\delta_g$ it is straightforward to calculate the variance of the estimated $\hat{\bar n}$:
      \begin{equation}
       {\rm Var}(\hat{\bar n})=\int \frac{\mathrm{d}{\bf l}^2}{(2\pi)^2}\left|\frac{W_{\bf l}}{\int \mathrm{d}\nv^2\,W(\nv)}\right|^2\,C_\ell.
      \end{equation}
      Here ${\bf l}$ is a 2D Fourier-space wave vector, $W_{\bf l}$ is the Fourier transform of the mask and $C_\ell$ is the power spectrum of $\delta_g$\footnote{We verified the validity of this calculation by replacing $C_\ell$ by its shot-noise contribution and recovering the Poisson limit (${\rm Var}_{\rm Poisson}(\hat{\bar n})=\bar{n}/A$, where $A$ is the footprint area). We find that the error on $\hat{\bar n}$ is dominated by cosmic variance (as opposed to shot noise) by more than a factor of $\sim5$ in all cases.}. Following this procedure, we estimate the number density in each field and redshift bin as well as its uncertainty (where we use the mask described in Section \ref{ssec:methods.mask} and the best-fit theory power spectra to compute the latter). The results are shown in Figure \ref{fig:ndens_consistency}. In all cases we find no significant deviations in the number density found in each field with respect to the mean, with only one estimate out of the 24 (GAMA15H in bin 2) deviating by more than 2 $\sigma$. We therefore conclude that there is no evidence of inconsistency between fields on the basis of their number densities.
      \begin{figure}
        \centering
        \includegraphics[width=0.7\textwidth]{figures/ndens_consistency.pdf}
        \caption{Source number densities estimated in the 6 different fields ($x$-axis) and in the 4 different redshift bins (in each panel, from top to bottom). We find no evidence of inconsistency between the number density of sources found in each field.}
        \label{fig:ndens_consistency}
      \end{figure}

      \paragraph{Power spectra.} Figure \ref{fig:cls_consistency} shows the difference between the power spectra estimated in each field and the coadded power spectra normalized by the power spectrum errors for all auto- and cross-correlations. We observe a reasonable scatter with respect to the coadded spectra of up to $\sim3\sigma$, which does not immediately indicate any evidence for inconsistency between fields. As a more quantitative check for inconsistencies we carry out a $\chi^2$ analysis of this scatter. Let  $\Delta {\bf C}_f={\bf C}_f-{\bf C}_{\rm coadd}$ be the difference between the power spectra measured in field $f$ and the coadded ones. Using the notation of Section \ref{sssec:results.spectra.fid}, the covariance of $\Delta {\bf C}_f$ is given by:
      \begin{equation}
        {\sf Cov}_{\Delta_f}=\left(1-2\frac{A_f}{\sum_{f'}A_{f'}}\right){\sf Cov}_f+{\sf Cov}_{\rm coadd}.
      \end{equation}
      We can therefore quantify the significance of the power spectrum differences by computing the $\chi^2$:
      \begin{equation}
        \chi^2\equiv\Delta{\bf C}_f^T\cdot{\sf Cov}_{\Delta_f}^{-1}\cdot\Delta{\bf C}_f,
      \end{equation}
      and its probability to exceed (PTE) under the assumption that $\chi^2$ follows a ``chi-squared'' distribution with a number of degrees of freedom given by the size of $\Delta{\bf C}_f$. Doing so for all the individual auto- and cross-correlations, as well as for the combined data vector containing all of them simultaneously, we find no quantitative evidence of inconsistency between fields. All PTEs are larger than 8\%, with the vast majority of them lying above 30\%. We conclude that there is no evidence for systematic biases from the power spectrum measurements in different fields, and therefore it is safe to coadd them and use the coadded spectra to obtain model constraints. We observe that the HECTOMAP field exhibits some of the lowest PTEs in the analysis described above. Although none of these are cause for concern, we have omitted the measurements from HECTOMAP when obtaining the coadded power spectra for safety. Since this is by far the smallest field ($\sim5\,{\rm deg}^2$), the associated loss of sensitivity is negligible.
      \begin{figure}
        \centering
        \includegraphics[width=0.99\textwidth]{figures/cls_consistency.pdf}
        \caption{Difference between the power spectra measured in each field and the mean coadded spectra normalized by the 1$\sigma$ errors. The pair of numbers in the upper left corner of each panel corresponds to the indices of the bins being cross-correlated. We find no inconsistency between the different measurements.}
        \label{fig:cls_consistency}
      \end{figure}

@@

  \subsection{Covariance matrix}\label{ssec:results.covariance}
    We compute the covariance matrix of all measured power spectra analytically as described in Sec.~\ref{sssec:methods.theory.covar} and the resulting correlation matrix is shown in Fig.~\ref{fig:covmat}. In Fig.~\ref{fig:covariance-contributions}, we split the auto-covariances for the four redshift bins into the Gaussian, non-Gaussian and SSC contributions. As can be seen, the Gaussian contribution is dominant on large scales while the non-Gaussian part becomes important at small scales and low redshift. The high redshift bins are mostly insensitive to non-Gaussian contributions to the covariance. Finally, the SSC is subdominant in all cases. This is mainly due to the number density correction in Eq.~\ref{eq:ps-resp}, which significantly suppresses any SSC contributions to the total covariance.
    \begin{figure}
      \begin{center}
        \includegraphics[width=0.5\textwidth]{figures/covmat_HSC_dg_hoddg_hodxdg_hoddg_hod_trisp=4h3h2h_SSC=LINBIAS-resp_l=ells_eff_all_wcross_clfit=HOD-zevol_culled=True_coadd=area_G+NG+SSC.pdf}
        \caption{Correlation matrix for all auto- and cross-power spectra considered in our analysis, computed analytically as described in Sec.~\ref{sssec:methods.theory.covar}.}\label{fig:covmat}
      \end{center}
    \end{figure}

@@

      \begin{figure}
        \begin{center}
          \includegraphics[width=0.95\textwidth]{figures/contours-Mmin-M1_mPk=HOD_fix=alpha-fc-sigmaM_HOD=zevol_fit=pz-shifts-pz-widths+prior=0p2_fit=auto+cross_cosmo=const_HOD-param=zfid_clfit=HOD-zevol_cov=G+NG+SSC-vs-fit=auto-vs-cov=G-vs-cov=G+SSC.pdf}
          \caption{Comparison of our fiducial constraints on HOD parameters to those obtained from auto-power spectra alone, accounting only for Gaussian and Gaussian and SSC contributions to the covariance matrix. The inner (outer) contour shows the $68 \%$ c.l. ($95 \%$ c.l.).} \label{fig:constraints-cov=G+NG+SSC-vs-cov=G-vs-cov=G+SSC-vs-fit=auto}
        \end{center}
      \end{figure}

@@

  DA acknowledges support from the Beecroft trust and from the Science and Technology Facilities Council through an Ernest Rutherford Fellowship, grant reference ST/P004474/1.
  HA has been supported by the Rutgers Discovery Informatics Institute Fellowship of Excellence in Computational and Data Science (AY 2017-2020) and Rutgers University \& Bevier Dissertation Completion Fellowship (AY 2019-2020); HA also thanks the LSSTC Data Science Fellowship Program, which is funded by LSSTC, NSF Cybertraining Grant No. 1829740, the Brinson Foundation, and the Moore Foundation, as participation in the program has benefited this work. HA, AB, and EG were supported by the Department of Energy (grants DE-SC0011636 and DE-SC0010008).
  ZG is supported by a Rhodes Scholarship granted by the Rhodes Trust.
  RM is supported by the Department of Energy Cosmic Frontier program, grant DE-SC0010118.
  HM has been supported by the Jet Propulsion Laboratory, California Institute of Technology, under a contract with the National Aeronautics and Space Administration and Grant-in-Aid for Scientific Research from the JSPS Promotion of Science (No. 18H04350). 
  JAN us supported by grant DOE DE-SC0007914 from the U.S. Department of Energy Office of Science (Office of High Energy Physics).
  AN acknowledges support from National Science Foundation Grant No. 1814971.
  This manuscript has been authored by Fermi Research Alliance, LLC under Contract No. DE-AC02-07CH11359 with the U.S. Department of Energy, Office of Science, Office of High Energy Physics.
  The work by JS was supported by the U.S. Department of Energy award DE-SC0009920.
  This work was supported by the grant MDM-2015-0509 from the Spanish Ministry of Science and Innovation.
  This project was supported in part by the U.S. Department of Energy, Office of Science, Office of Workforce Development for Teachers and Scientists (WDTS) under the Science Undergraduate Laboratory Internships Program (SULI).

>>>>

  \begin{table}
  \centering
  \begin{tabular}{|l|l|l|l|}
    \hline
    $z_{\rm phot}^{\rm ini}$ & $z_{\rm phot}^{\rm end}$ & $\bar{z}$ & $N_{\rm gal}$ \\
    \hline
    0.15 & 0.5  & 0.57 & 1,750,274 \\
    0.5  & 0.75 & 0.68 & 1,766,939 \\
    0.75 & 1.0  & 0.91 & 1,702,685 \\
    1.0  & 1.5  & 1.26 & 1,752,359 \\
    \hline
  \end{tabular}
  \caption{Summary of the 4 redshift bins used in our analysis. The first two columns show the photo-$z$ bin edges. We used the {\tt photoz\_best} redshift estimator for the photo-$z$ code {\tt Ephor\_AB} to assign galaxies to different bins. The third column shows the mean redshift of each bin calculated from the fiducial redshift distributions described in Section \ref{ssec:methods.nz}. The last column shows the number of galaxies in each bin. The bin edges were chosen to roughly contain an equal number of galaxies in each of them.} \label{tab:bins_summary}
  \end{table}

@@

  \subsection{Redshift distributions}\label{ssec:methods.nz}
    \begin{figure}
      \centering
      \includegraphics[width=0.9\textwidth]{figures/nzs.pdf}
      \caption{Redshift distributions used in this analysis. The histograms in each panel are the same, and correspond to the redshift distributions estimated from the COSMOS 30-band catalog. These are the fiducial redshift distributions used in our analysis. The solid curves in each panel show the stacked redshift distributions inferred by stacking the per-object photo-$z$ probability distributions obtained with the four alternative photo-$z$ codes explored here ({\tt DEmP}, {\tt Ephor}, {\tt Ephor\_AB} and {\tt FRANKEN-Z} from left to right and top to bottom). The redshift distributions for the four different redshift bins listed in Table \ref{tab:bins_summary} are shown in different colors. Note that all four redshift estimation codes use the same COSMOS 30-band catalog so the level of visual agreement might not provide a realistic measure of uncertainty on the redshift distributions.}
      \label{fig:nzs}
    \end{figure}
    The underlying redshift distributions of our tomographic samples ($p^i(z)$ in Eq. \ref{eq:cell_gg_limber}) are a central component of the theory model used to infer astrophysical and cosmological parameters. Estimating redshift distributions for photometric samples is a non-trivial problem that has been studied extensively in the literature \citep{2008MNRAS.390..118L,2008ApJ...684...88N,2018MNRAS.478..592H}. We use two different methods to estimate $p^i(z)$.

@@

      \paragraph{Number densities.} The number density of galaxies estimated in a given field is simply given by the ratio of the number of galaxies and the area of the field:
      \begin{equation}
        \hat{\bar{n}}=\frac{\sum_p N_p}{\sum_p \Omega_p}= \bar{n} \frac{\int \mathrm{d}\nv^2\,W(\nv)\left[1+\delta_g(\nv)\right]}{\int \mathrm{d}\nv^2\,W(\nv)},
      \end{equation}
      where $p$ runs over all pixels in the map, and $N_p$ and $\Omega_p$ are the number of objects and area of pixel $p$. In the second equality we have taken the continuum limit, $\bar{n}$ is the true number density, $\delta_g$ is the galaxy overdensity and $W$ is the field's mask. Using the statistics of $\delta_g$ it is straightforward to calculate the variance of the estimated $\hat{\bar n}$:
      \begin{equation}
       {\rm Var}(\hat{\bar n})=\int \frac{\mathrm{d}{\bf l}^2}{(2\pi)^2}\left|\frac{W_{\bf l}}{\int \mathrm{d}\nv^2\,W(\nv)}\right|^2\,C_\ell.
      \end{equation}
      Here ${\bf l}$ is a 2D Fourier-space wave vector, $W_{\bf l}$ is the Fourier transform of the mask and $C_\ell$ is the power spectrum of $\delta_g$\footnote{We verified the validity of this calculation by replacing $C_\ell$ by its shot-noise contribution and recovering the Poisson limit (${\rm Var}_{\rm Poisson}(\hat{\bar n})=\bar{n}/A$, where $A$ is the footprint area). We find that the error on $\hat{\bar n}$ is dominated by cosmic variance (as opposed to shot noise) by more than a factor of $\sim5$ in all cases.}. Following this procedure, we estimate the number density in each field and redshift bin as well as its uncertainty (where we use the mask described in Section \ref{ssec:methods.mask} and the best-fit theory power spectra to compute the latter). The results are shown in Figure \ref{fig:ndens_consistency}. In all cases we find no significant deviations in the number density found in each field with respect to the mean, with only one estimate out of the 24 (GAMA15H in bin 2) deviating by more than 2 $\sigma$. We therefore conclude that there is no evidence of inconsistency between fields on the basis of their number densities.
      \begin{figure}
        \centering
        \includegraphics[width=0.7\textwidth]{figures/ndens_consistency.pdf}
        \caption{Source number densities estimated in the 6 different fields ($x$-axis) and in the 4 different redshift bins (in each panel, from top to bottom). We find no significant evidence of inconsistency between the number density of sources found in each field.}
        \label{fig:ndens_consistency}
      \end{figure}

      \paragraph{Power spectra.} Figure \ref{fig:cls_consistency} shows the difference between the power spectra estimated in each field and the coadded power spectra normalized by the power spectrum errors for all auto- and cross-correlations. We observe a reasonable scatter with respect to the coadded spectra of up to $\sim3\sigma$, which does not immediately indicate any evidence for inconsistency between fields. As a more quantitative check for inconsistencies we carry out a $\chi^2$ analysis of this scatter. Let  $\Delta {\bf C}_f={\bf C}_f-{\bf C}_{\rm coadd}$ be the difference between the power spectra measured in field $f$ and the coadded ones. Using the notation of Section \ref{sssec:results.spectra.fid}, the covariance of $\Delta {\bf C}_f$ is given by:
      \begin{equation}
        {\sf Cov}_{\Delta_f}=\left(1-2\frac{A_f}{\sum_{f'}A_{f'}}\right){\sf Cov}_f+{\sf Cov}_{\rm coadd}.
      \end{equation}
      We can therefore quantify the significance of the power spectrum differences by computing the $\chi^2$:
      \begin{equation}
        \chi^2\equiv\Delta{\bf C}_f^T\cdot{\sf Cov}_{\Delta_f}^{-1}\cdot\Delta{\bf C}_f,
      \end{equation}
      and its probability to exceed (PTE) under the assumption that $\chi^2$ follows a ``chi-squared'' distribution with a number of degrees of freedom given by the size of $\Delta{\bf C}_f$. Doing so for all the individual auto- and cross-correlations, as well as for the combined data vector containing all of them simultaneously, we find no quantitative evidence of inconsistency between fields. All PTEs are larger than 8\%, with the vast majority of them lying above 30\%. We conclude that there is no evidence for systematic biases from the power spectrum measurements in different fields, and therefore it is safe to coadd them and use the coadded spectra to obtain model constraints. We observe that the HECTOMAP field exhibits some of the lowest PTEs in the analysis described above. Although none of these are cause for concern, we have omitted the measurements from HECTOMAP when obtaining the coadded power spectra for safety. Since this is by far the smallest field ($\sim5\,{\rm deg}^2$), the associated loss of sensitivity is negligible.
      \begin{figure}
        \centering
        \includegraphics[width=0.99\textwidth]{figures/cls_consistency.pdf}
        \caption{Difference between the power spectra measured in each field and the mean coadded spectra normalized by the 1$\sigma$ errors. The pair of numbers in the upper left corner of each panel corresponds to the indices of the bins being cross-correlated. We find consistency between the different measurements.}
        \label{fig:cls_consistency}
      \end{figure}

@@

  \subsection{Covariance matrix}\label{ssec:results.covariance}
    We compute the covariance matrix of all measured power spectra analytically as described in Sec.~\ref{sssec:methods.theory.covar} and the resulting correlation matrix\footnote{The correlation matrix $\mathsf{Corr}$ is obtained from the covariance matrix $\mathsf{C}$ as $\mathsf{Corr}_{ij} = \sfrac{\mathsf{C}_{ij}}{\sqrt{\mathsf{C}_{ii} \mathsf{C}_{jj}}}$.} is shown in Fig.~\ref{fig:covmat}. In Fig.~\ref{fig:covariance-contributions}, we split the auto-covariances for the four redshift bins into the Gaussian, non-Gaussian and SSC contributions. As can be seen, the Gaussian contribution is dominant on large scales while the non-Gaussian part becomes important at small scales and low redshift. The high redshift bins are mostly insensitive to non-Gaussian contributions to the covariance. Finally, the SSC is subdominant in all cases. This is mainly due to the number density correction in Eq.~\ref{eq:ps-resp}, which significantly suppresses any SSC contributions to the total covariance.
    \begin{figure}
      \begin{center}
        \includegraphics[width=0.5\textwidth]{figures/covmat_HSC_dg_hoddg_hodxdg_hoddg_hod_trisp=4h3h2h_SSC=LINBIAS-resp_l=ells_eff_all_wcross_clfit=HOD-zevol_culled=True_coadd=area_G+NG+SSC.pdf}
        \caption{Correlation matrix for all auto- and cross-power spectra considered in our analysis, computed analytically as described in Sec.~\ref{sssec:methods.theory.covar}.}\label{fig:covmat}
      \end{center}
    \end{figure}

@@

      \begin{figure}
        \begin{center}
          \includegraphics[width=0.95\textwidth]{figures/contours-Mmin-M1_mPk=HOD_fix=alpha-fc-sigmaM_HOD=zevol_fit=pz-shifts-pz-widths+prior=0p2_fit=auto+cross_cosmo=const_HOD-param=zfid_clfit=HOD-zevol_cov=G+NG+SSC-vs-fit=auto-vs-cov=G-vs-cov=G+SSC.pdf}
          \caption{Comparison of our fiducial constraints on HOD parameters to those obtained from auto-power spectra alone, accounting only for Gaussian or Gaussian and SSC contributions to the covariance matrix. The inner (outer) contour shows the $68 \%$ c.l. ($95 \%$ c.l.).} \label{fig:constraints-cov=G+NG+SSC-vs-cov=G-vs-cov=G+SSC-vs-fit=auto}
        \end{center}
      \end{figure}

@@

  DA acknowledges support from the Beecroft trust and from the Science and Technology Facilities Council through an Ernest Rutherford Fellowship, grant reference ST/P004474/1.
  HA has been supported by the Rutgers Discovery Informatics Institute Fellowship of Excellence in Computational and Data Science (AY 2017-2020) and Rutgers University \& Bevier Dissertation Completion Fellowship (AY 2019-2020); HA also thanks the LSSTC Data Science Fellowship Program, which is funded by LSSTC, NSF Cybertraining Grant No. 1829740, the Brinson Foundation, and the Moore Foundation, as participation in the program has benefited this work. HA, AB, and EG were supported by the Department of Energy (grants DE-SC0011636 and DE-SC0010008).
  ZG is supported by a Rhodes Scholarship granted by the Rhodes Trust.
  RM is supported by the Department of Energy Cosmic Frontier program, grant DE-SC0010118.
  HM has been supported by the Jet Propulsion Laboratory, California Institute of Technology, under a contract with the National Aeronautics and Space Administration and Grant-in-Aid for Scientific Research from the JSPS Promotion of Science (No. 18H04350). 
  JAN us supported by grant DOE DE-SC0007914 from the U.S. Department of Energy Office of Science (Office of High Energy Physics).
  JD and AN acknowledge support from National Science Foundation Grant No. 1814971.
  This manuscript has been authored by Fermi Research Alliance, LLC under Contract No. DE-AC02-07CH11359 with the U.S. Department of Energy, Office of Science, Office of High Energy Physics.
  The work by JS was supported by the U.S. Department of Energy award DE-SC0009920.
  This work was supported by the grant MDM-2015-0509 from the Spanish Ministry of Science and Innovation.
  This project was supported in part by the U.S. Department of Energy, Office of Science, Office of Workforce Development for Teachers and Scientists (WDTS) under the Science Undergraduate Laboratory Internships Program (SULI).
